---
title: "Pok&eacute;monGo Machine Learning Report"
author: "Hunter Campbell"
date: "June 20, 2017"
output: rmarkdown::github_document
---

```{r echo = FALSE, message = FALSE}
library(ggplot2)
library(dplyr)
library(plyr)
library(plotly)
library(tidyr)
PokeGo_clean <- read.csv(file = "C:/Users/hcnur_000/Desktop/Capstone Project/Capstone Project Data Wrangling/PokeGo_clean.csv")
PokeGo_water2water <- read.csv(file = "C:/Users/hcnur_000/Desktop/Capstone Project/Capstone Project Data Wrangling/PokeGo_water2water.csv")
PokeGo_notwater2water <- read.csv(file = "C:/Users/hcnur_000/Desktop/Capstone Project/Capstone Project Data Wrangling/PokeGo_notwater2water.csv")
PokeGo_clean <- select(PokeGo_clean, -X)
PokeGo_water2water <- select(PokeGo_water2water, -X)
PokeGo_notwater2water <- select(PokeGo_notwater2water, -X)
```

<br>

###**The Question**
The machine learning question for *Pok&eacute;monGo* is: **How to calculate Pok&eacute;mon rarity**.

If Pok&eacute;mon rarity can be calculated, then a *Pok&eacute;monGo* user would know what Pok&eacute;mon are truly rare (and worth searching for), and which are common. That's what will be solved in this report.

<br>

###**The Machine Learning Technique**
A supervised approach was used to solve Pok&eacute;mon rarity. To be specific, the K-means algorithm was used on the independent variable of Pok&eacute;mon Id. This way, rarity could be clustered into a few different groups. Four groups were chosen: Common, Uncommon, Rare, and Very Rare. So, four clusters were made. First, the data had to be set-up to be able to make these clusters.

Here's the code for the set-up:

```{r}
#First I made new datasets:
cleanId <- select(PokeGo_clean, pokemonId)
waterId <- select(PokeGo_water2water, pokemonId)
notwaterId <- select(PokeGo_notwater2water, pokemonId)

#I turned them into a factor, because each Id represents a Pokemon.
cleanIdFactor <- as.factor(cleanId$pokemonId)
waterIdFactor <- as.factor(waterId$pokemonId)
notwaterIdFactor <- as.factor(notwaterId$pokemonId)

#I then counted the number of times each Pokemon is seen:
cleanId <- count(cleanIdFactor, 1)
waterId <- count(waterIdFactor, 1)
notwaterId <- count(notwaterIdFactor, 1)

#and arranged the datasets by the frequency (this way k-means will group it
#according to the rarity/frequency instead of the Id).
cleanId <- arrange(cleanId, freq)
waterId <- arrange(waterId, freq)
notwaterId <- arrange(notwaterId, freq)
#I then changed the x columns name back to PokemonId, because count deleted the 
#old PokemonId column and replaced it with a column named x.
colnames(cleanId)[1] <- "pokemonId"
colnames(waterId)[1] <- "pokemonId"
colnames(notwaterId)[1] <- "pokemonId"
```

Now that the code is set-up for K-means Clustering, the clusters can be made.

Here's the code:

```{r}
k <- 4
set.seed(1)

#Making the clusters for each dataset:
cleanIdKMC <- kmeans(cleanId$freq, centers = k)
waterIdKMC <- kmeans(waterId$freq, centers = k)
notwaterIdKMC <- kmeans(notwaterId$freq, centers = k)
```

<br>

###**The Results**
Now that each dataset has been clustered, it can now be evaluated to show the results.

**All Pok&eacute;mon**

```{r echo = FALSE, message = FALSE, fig.height = 3}
cleanRarity <- ggplot(cleanId, aes(pokemonId, freq, col = factor(cleanIdKMC$cluster), alpha = .5)) +
  geom_point() +
  labs(x = "All Pokemon Id's", y = "Frequency") +
  scale_color_manual(values = c("#800080", "#006622", "#e68a00", "#ff0000"),
                     name = "Rarity",
                     breaks = c(4, 3, 2, 1),
                     labels = c("Very Common", "Common", "Rare", "Very Rare")) +
  scale_alpha(guide = F) +
  theme(axis.text.x = element_blank())
cleanRarity

#Can't use this in github...
#ggplotly(cleanRarity) %>%
#  layout(showlegend = F)
```

<br>

**Water Type Pok&eacute;mon**

```{r echo = FALSE, message = FALSE, fig.height = 3}
waterRarity <- ggplot(waterId, aes(pokemonId, freq, col = factor(waterIdKMC$cluster), alpha = .5)) +
  geom_point() +
  labs(x = "Water Type Pokemon Id's", y = "Frequency") +
  scale_color_manual(values = c("#800080", "#e68a00", "#ff0000", "#006622"),
                     name = "Rarity",
                     breaks = c(4, 3, 2, 1),
                     labels = c("Very Common", "Common", "Rare", "Very Rare")) +
  scale_alpha(guide = F) +
  theme(axis.text.x = element_blank())
waterRarity

#Can't use this in github...
#ggplotly(waterRarity) %>%
#  layout(showlegend = F)
```

<br>

**Non-water Type Pok&eacute;mon**

```{r echo = FALSE, message = FALSE, fig.height = 3}
notwaterRarity <- ggplot(notwaterId, aes(pokemonId, freq, col = factor(notwaterIdKMC$cluster), alpha = .5)) +
  geom_point() +
  labs(x = "Non-water Type Pokemon Id's", y = "Frequency") +
  scale_color_manual(values = c("#ff0000", "#800080", "#e68a00", "#006622"),
                     name = "Rarity",
                     breaks = c(4, 3, 2, 1),
                     labels = c("Very Common", "Common", "Rare", "Very Rare")) +
  scale_alpha(guide = F) +
  theme(axis.text.x = element_blank())
notwaterRarity

#Can't use this in github...
#ggplotly(notwaterRarity) %>%
#  layout(showlegend = F)
```

<br>

###**Conclusion**
As each graph demonstrates, the K-means Algorithm calculated rarity under each case. It even calculated the rarity of water type Pok&eacute;mon; which means it can calculate any specific Pok&eacute;mon type, or even a group of Pok&eacute;mon (like the non-water type graph). In this way, any *Pok&eacute;monGo* user can know which Pok&eacute;mon are rare, and which are common.

<br>




