---
title: "Pok&eacute;monGo Milestone Report"
author: "Hunter Campbell"
date: "June 16, 2017"
output: rmarkdown::github_document
---

```{r echo = FALSE, message = FALSE}
library(ggplot2)
library(dplyr)
PokeGo_clean <- read.csv(file = "C:/Users/hcnur_000/Desktop/Capstone Project/Capstone Project Data Wrangling/PokeGo_clean.csv")
PokeGo_water2water <- read.csv(file = "C:/Users/hcnur_000/Desktop/Capstone Project/Capstone Project Data Wrangling/PokeGo_water2water.csv")
PokeGo_notwater2water <- read.csv(file = "C:/Users/hcnur_000/Desktop/Capstone Project/Capstone Project Data Wrangling/PokeGo_notwater2water.csv")
PokeGo_clean <- select(PokeGo_clean, -X)
PokeGo_water2water <- select(PokeGo_water2water, -X)
PokeGo_notwater2water <- select(PokeGo_notwater2water, -X)
```

<br>

###**Introduction**
*Pok&eacute;monGo* is an augmented reality game for handheld devices created by *Niantic, Inc.* With over 100 million downloads, *Pok&eacute;monGo* is one of the most popular apps ever created. This report shows the correlation between Pok&eacute;mon spawn frequency and different environments that a *Pok&eacute;monGo* user might encounter. This report also presents data and graphical analysis that will affirm a rise in rare Pok&eacute;mon spawn rates and a drop in common Pok&eacute;mon spawn rates--- all based on location. This report ultimately can help all *Pok&eacute;monGo* users with their quest to "Catch'em All". Lets get started!

<br>

###**The Data Wrangling**
The original data is too big to upload to github, so here is a link to download the data: [Original Pok&eacute;monGo Data](https://www.kaggle.com/semioniy/predictemall)

There were several columns from the original data that didn't make sense to include in this report. This shows the columns that were picked and how they were written in *RStudio*:

```{r eval = FALSE}
PokeGo_clean <- select(PokeGo_original, pokemonId, appearedTimeOfDay,
                       appearedDayOfWeek, terrainType:closeToWater,
                       weather, population_density,
                       gymDistanceKm, pokestopDistanceKm)
```

The new data (clean) was then ordered according to our independent variable (Pokemon Id):

```{r eval = FALSE}
library("dplyr")
PokeGo_clean <- PokeGo_clean %>% arrange(pokemonId)
```

Afterwards, any blank and/or NA values were checked for (both returned as zero):

```{r eval = FALSE}
sum(is.na(PokeGo_clean))
sum(PokeGo_clean == "")
```

In the Day of Week column, Monday wasn't found, but there was a variable listed as "dummy_variable". The dummy variable was then changed to Monday:

```{r eval = FALSE}
PokeGo_clean$appearedDayOfWeek <- gsub("dummy_day", "Monday", PokeGo_clean$appearedDayOfWeek, ignore.case = T)
```

The gym and pok&eacute;stop distance columns were originally labelled as kilometers, which was hard to read, because of the decimal placement. They were both changed to read as meters:

```{r eval = FALSE}
PokeGo_clean$gymDistanceKm <- PokeGo_clean$gymDistanceKm * 1000
PokeGo_clean$pokestopDistanceKm <- as.numeric(PokeGo_clean$pokestopDistanceKm) * 1000
#NA's were created in the pokestop distance column, here's the fix (replaced with the mean of the column):
PokeGo_clean$pokestopDistanceKm[is.na(PokeGo_clean$pokestopDistanceKm)] <- mean(PokeGo_clean$pokestopDistanceKm, na.rm = T)

#Changing the titles of the two columns:
colnames(PokeGo_clean)[8] <- "gymDistance_m"
colnames(PokeGo_clean)[9] <- "pokestopDistance_m"
```

After creating the clean data, two new datasets were created to later test the correlation of Pok&eacute;mon to Close to Water. This is how they were made:

```{r eval = FALSE}
#The water type Pokemon dataset:
PokeGo_water2water <- filter(PokeGo_clean, pokemonId ==
                               c(7:9, 54:55, 60:62, 72:73, 79:80,
                                 86:87, 90:91, 98:99, 116:121,
                                 129:131, 134, 138:141)) %>%
    select(pokemonId, closeToWater)

#The non-water type Pokemon dataset:
PokeGo_notwater2water <- filter(PokeGo_clean, pokemonId ==
                                  c(1:6, 10:53, 56:59, 63:71, 74:78,
                                    81:85, 88:89, 92:97, 100:115,
                                    122:128, 132:133, 135:137, 142:151)) %>%
    select(pokemonId, closeToWater)
```

<br>

###**The Data**
These are the top few rows of the data that will be used (which was made in the Data Wrangling section):

**The Clean Data**
  
```{r echo = FALSE}
head(PokeGo_clean)
```

<br>

**Water Type Pok&eacute;mon Data**

```{r echo = FALSE}
head(PokeGo_water2water)
```

<br>

**Non-water Type Pok&eacute;mon Data**

```{r echo = FALSE}
head(PokeGo_notwater2water)
```

<br>

###***Pok&eacute;monGo* Rarity Chart**
To start us off, we need to know the difference between the Pok&eacute;mon that are common, and the Pok&eacute;mon that are more rare. The rarity in *Pok&eacute;monGo* is different than the older *Gameboy* to *DS* games. In *Pok&eacute;monGo*, the rarity is based on how often a Pok&eacute;mon spawns. As the graph below shows, there is a peak of Pok&eacute;mon Id's from 10-25 (the common Pok&eacute;mon).  There are also a few shorter peaks around 40-50, 95-100, and 130-135 (the less common Pok&eacute;mon). The rest of the graph shows the rare Pok&eacute;mon.

Here's a reference on what each Pok&eacute;mon Id is: [Pok&eacute;mon Id's](https://bulbapedia.bulbagarden.net/wiki/List_of_Pok%C3%A9mon_by_Kanto_Pok%C3%A9dex_number)

```{r echo = FALSE, fig.height = 3, fig.align = "center"}
ggplot(PokeGo_clean, aes(pokemonId)) +
  geom_bar(fill = "red") +
  labs(x = "Pokemon Id", y = "Pokemon Appearances")
```

<br>

###**Pok&eacute;mon's Preferred Circumstances**
These four graphs show different Pok&eacute;mon's preferred circumstances to spawn during. 

The first graph shows what days of the week Pok&eacute;mon usually spawn. The most popular days are Wednesday through Saturday, with the peak being Friday and Saturday.

The second graph shows what time Pok&eacute;mon most often spawn during. Night time is, interestingly, the most popular time to spawn, while the other times of day are around the same height.

The third graph shows the preferred weather that Pok&eacute;mon spawn during.

The fourth graph shows the type of terrain most Pok&eacute;mon spawn in. This most likely is just the terrain type of where users live, but that would also mean that there is a common ground for what terrain Pok&eacute;mon spawn.

We can assume from the graphs that there is a correlation between when and where Pok&eacute;mon spawn and a *Pok&eacute;monGo* user's activity with the app. This would mean that *Pok&eacute;monGo* users usually play on the weekend, during good weather, in a common location, and/or during the night. These graphs do not present anything further than this concerning the why (e.g. Why during the night?) for each of these items.

Here's a reference for what each number equals for each terrain: [Terrain Types](http://glcf.umd.edu/data/lc/)

```{r echo = FALSE, fig.height = 3, fig.width = 4.5}
ggplot(PokeGo_clean, aes(appearedDayOfWeek)) +
  geom_bar(fill = "steelblue") +
  scale_x_discrete(limits = c("Monday", "Tuesday", "Wednesday", "Thursday",
                              "Friday", "Saturday", "Sunday")) +
  labs(x = "Day of Week", y = "Pokemon Appearances")

ggplot(PokeGo_clean, aes(appearedTimeOfDay)) +
  geom_bar(fill = "steelblue") +
  labs(x = "Time of Day", y = "Pokemon Appearances")

ggplot(PokeGo_clean, aes(weather)) +
  geom_bar(fill = "steelblue") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .25)) +
  labs(x = "Weather", y = "Pokemon Appearances")

ggplot(PokeGo_clean, aes(terrainType)) +
  geom_bar(fill = "steelblue") +
  labs(x = "Terrain Type", y = "Pokemon Appearances")
```

<br>

###**Pok&eacute;mon Spawn Frequency Correlations**
**Pok&eacute;mon to Close to Water**

The first graph is the control for the other two graphs, because it contains every Pok&eacute;mon.

The second graph shows the correlation between water type Pok&eacute;mon near water, and the graph after that shows the correlation between non-water type Pok&eacute;mon near water. Be mindful of the numbers on the y-axis, because this makes a difference in how our graphs are analyzed.

```{r echo = FALSE, fig.height = 3, fig.width = 3}
ggplot(PokeGo_clean, aes(closeToWater, fill = closeToWater)) +
  geom_bar() +
  labs(x = "Close To Water", y = "Appearances of All Pokemon Types", fill = "Close To Water")

ggplot(PokeGo_water2water, aes(closeToWater, fill = closeToWater)) +
  geom_bar() +
  labs(x = "Close to Water", y = "Appearances of Water Type Pokemon",
       fill = "Close To Water")

ggplot(PokeGo_notwater2water, aes(closeToWater, fill = closeToWater)) +
  geom_bar() +
  labs(x = "Close to Water", y = "Appearances of Non-water Type Pokemon",
       fill = "Close To Water")
```

There is a definite higher probability of finding a water type Pok&eacute;mon near water vs. a non-water type Pok&eacute;mon near water. Lets calculate this:

```{r}
summary(PokeGo_water2water$closeToWater)

#To find the percentage:
13901/(24507+13901)
```

This returns about a 36.2% chance of seeing a water type Pok&eacute;mon near water. Now lets calculate the chance of seeing a non-water type Pok&eacute;mon near water:

```{r}
summary(PokeGo_notwater2water$closeToWater)

#To find the percentage:
54598/(203015+54598)
```

This returns about a 21.2% chance of seeing a non-water type Pok&eacute;mon near water. That's a 15% difference! There is a 15% higher probability of finding a water type Pok&eacute;mon near water.

<br>

**Pok&eacute;mon to Population Density**

The first is a density plot. It makes sense that our density is mostly around the mean of 1,313.6. This code line shows the mean of population density:

```{r}
summary(PokeGo_clean$population_density)
```

The first graph also shows similar peaks to the Pok&eacute;mon Id graph. (This will be compared towards the end of the Pok&eacute;mon Spawn Frequency Correlations section.)

The second graph is a visualization of the rarity of Pok&eacute;mon in a densely populated area. It's interesting that the peaks and valleys are quite opposite to the Pok&eacute;mon Id graph, except at the end, where there is a constant decline. This shows that a *Pok&eacute;monGo* user has a higher chance of seeing rare Pok&eacute;mon while in densly populated areas.

```{r echo = FALSE, message = FALSE, fig.height = 3, fig.width = 4.5}
ggplot(PokeGo_clean, aes(pokemonId, population_density)) +
  geom_point(shape = 1, alpha = .1) +
  geom_density2d(color = "red", size = 1) +
  labs(x = "Pokemon Id", y = "Population Density")

ggplot(PokeGo_clean, aes(pokemonId, population_density)) +
  geom_smooth() +
  labs(x = "Pokemon Id", y = "Population Density")
```

<br>

**Pok&eacute;mon to Gym Distance**

The first graph demonstrates that there is a higher density of Pok&eacute;mon near gyms, because the closer the graph approaches zero on the y-axis, the closer a *Pok&eacute;monGo* user is to a gym.

The second graph then shows that there is a rise in seeing rare Pok&eacute;mon near gyms and a decline of common Pok&eacute;mon near gyms. Being by a gym would lead to greater success finding rare Pok&eacute;mon. There is some error ribbons to be noted, with great error around the Pok&eacute;mon Id's of 130-150, but the line of best fit shows great results.

```{r echo = FALSE, message = FALSE, fig.height = 3, fig.width = 4.5}
ggplot(PokeGo_clean, aes(pokemonId, gymDistance_m)) +
  geom_point(shape = 1, alpha = .1) +
  coord_cartesian(ylim = c(0, 1e+05)) +
  labs(x = "Pokemon Id", y = "Gym Distance in Meters")

ggplot(PokeGo_clean, aes(pokemonId, gymDistance_m)) +
  geom_smooth() +
  labs(x = "Pokemon Id", y = "Gym Distance in Meters")
```

<br>

**Pok&eacute;mon to Pok&eacute;stop Distance**

Again, the first graph shows that there is a rise in Pok&eacute;mon spawn rates the closer a *Pok&eacute;monGo* user is to a pok&eacute;stop. Being near a pok&eacute;stop is more demanding than being near a gym, though, because the density plots don't reach quite as high as the gym density plot.

Interestingly, the second graph shows a steady incline for rare Pok&eacute;mon with Id's between 125-150. This shows similar results to the second gym graph, which means that there is a higher chance of finding rare Pok&eacute;mon near pok&eacute;stops! Again, the error ribbons need to be noted, with greater error around the Pok&eacute;mon Id's of 130-150. Also, the numbers on the y-axis for this graph are much lower than the gym graph, which again clarifies that pok&eacute;stops are much more demanding on how close a *Pok&eacute;monGo* user needs to be to the pok&eacute;stop to see results.

```{r echo = FALSE, message = FALSE, fig.height = 3, fig.width = 4.5}
ggplot(PokeGo_clean, aes(pokemonId, pokestopDistance_m)) +
  geom_point(shape = 1, alpha = .1) +
  coord_cartesian(ylim = c(0, 1e+05)) +
  labs(x = "Pokemon Id", y = "Pokestop Distance in Meters")

ggplot(PokeGo_clean, aes(pokemonId, pokestopDistance_m)) +
  geom_smooth() +
  labs(x = "Pokemon Id", y = "Pokestop Distance in Meters")
```

<br>

**Pok&eacute;mon Id to Pok&eacute;mon Rarity**

Here's a comparison of the three line graphs, showing Pok&eacute;mon rarity, next to the Pok&eacute;mon Id graph:

```{r echo = FALSE, message = FALSE, fig.height = 3, fig.width = 4.5}
ggplot(PokeGo_clean, aes(pokemonId)) +
  geom_bar(fill = "red") +
  labs(x = "Pokemon Id", y = "Pokemon Appearances")

ggplot(PokeGo_clean, aes(pokemonId, population_density)) +
  geom_smooth() +
  labs(x = "Pokemon Id", y = "Population Density")

ggplot(PokeGo_clean, aes(pokemonId, gymDistance_m)) +
  geom_smooth() +
  labs(x = "Pokemon Id", y = "Gym Distance in Meters")

ggplot(PokeGo_clean, aes(pokemonId, pokestopDistance_m)) +
  geom_smooth() +
  labs(x = "Pokemon Id", y = "Pokestop Distance in Meters")
```

<br>

###**In Conlusion**
To wrap everything up, the graphs clearly show a relationship between Pok&eacute;mon in *Pok&eacute;monGo* and the environments in which they spawn most often. There is a higher chance of seeing a water type Pok&eacute;mon near water, and a higher chance of seeing rare Pok&eacute;mon in densly populated areas, near gyms, and near pok&eacute;stops. This can be useful to any *Pok&eacute;monGo* user, especially if they are on the hunt for rare Pok&eacute;mon.

<br>
